---
title: "Covid91 vaccine study Final2023"
author: "Matt Diaz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: " April 3, 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_o$: There is no difference in the effectiveness of the covid 921 vaccine in different subgroups.

$h_a$: There is a difference in the effectiveness of the civid 921 vaccine in different subgroups.

## Methods

The variables in this study: Males-categorical, Females-categorical, LGBTQ-categorical, Druggies, categorical. We will be using the Cat~Cat helper file. 

## Graphical Descriptive Results

#Females

```{r}
dataF = filter(FinalData,sex=="F")
barchartGC(~infected + treatment,data=dataF)
```

```{r}
barchartGC(~infected + treatment, data=dataF, type= "percent")
```

This graph shows us that of all the females that have contracted COVID, the females who received the vaccine had less symptoms than those who were administered the placebo, while the all the female group that did not contract COVID showed no difference between those who were vaccinated and not vaccinated. 


### Numerical Descriptive

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=dataF)
rowPerc(table1)
```

```{r}
colPerc(table1)
```

With the data presented, it seems as though the vaccine has a lower infection rate compared to the placebo. In addition, among those infected, the vaccine has a slightly better treatment outcome with there being a .67 difference between the placebo and vaccine. The data from this table shows that from the female group that were infected with COVID 19, those who were unvaccinated, 60% of them experienced symptoms while only 40% of the female group who were vaccinated experienced symptoms. In addition, this also shows that within the female placebo group, 2.04% were infected with COVID, while only 1.37% of the vaccinated group contracted COVID.


## Inferential Results

```{r}
chisq.test(table1)
```


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value is 0.0002732 for the Chi-squared test, that there is a significant association between infected and treatment within the Female subroup.

Based on the Fisher exact test, we got a p-value of 0.0002162, which suggest the same as the Chi-squared test.


# LGBTQ

## Introduction

$H_o$: There is no difference in effectiveness of the Covid vaccine between LGBTQ people.

$h_a$: There is a difference in effectiveness of the Covid vaccine between LGBTQ people.

## Methods

We will be using the Cat~Cat helper file. 

## Graphical Descriptive Results

```{r}
levels(FinalData$LGBTQ)
```

```{r}
dataG = filter(FinalData,LGBTQ=="gay")
barchartGC(~infected + treatment,data=dataG)
```

```{r}
barchartGC(~infected + treatment,data=dataG, type = "percent")
```

This chart shows that there is a significant difference between the placebo and the vaccine. Among those that were infected with covid, those who had the vaccine were infected more than those who had received the placebo. This graph is showing that within the LGBTQ subgroup, those who contracted the COVID virus and received the vaccination experienced significantly more symptoms than those who did not receive the vaccine. Those who did not get COVID and were both vaccinated and unvaccinated did not show any difference.


## Numerical Results 

```{r}
table1 <-xtabs(~infected + treatment, data=dataG)
rowPerc(table1)
```

```{r}
colPerc(table1)
```

This table is showing us that within the LGBTQ placebo group, those individuals experienced 15.38% of symptoms while the vaccinated group experienced 84.62% of symptoms. This graph also shows us that 0.66% of the LGBTQ group who had the placebo contracted COVID, while 3.54% of the vaccinated group contracted COVID.


## Inferential Results

```{r}
chisq.test(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

With the p-value being zero, we reject the null hypothesis and therefore conclude that there is a significant association between the treatment type and infection status. The treatment type, whether that be someone who was administered the vaccine or received the placebo, is related to getting infected with the virus. The chisq test shows that the p-value for the LGBTQ group is 0.0000006668 showing that there is an effect between the COVID vaccine and the LGBTQ subgroup, concluding that we are rejecting the null hypothesis.  


#Druggies

## Introduction

$H_o$: There is no difference in the effectiveness of COVID921 vaccine within druggies.

$h_a$:There is a difference in the effectiveness of COVID921 vaccine within druggies.

## Methods

We will be using the Cat~Cat helper file.

## Graphical Descriptive Results

```{r}
levels(FinalData$DrugUser)
```

```{r}
dataD = filter(FinalData,DrugUser=="yes")
barchartGC(~infected + treatment,data=dataD)
```

```{r}
barchartGC(~infected + treatment,data=dataD, type="percent")
```

This graph us showing that from the drug users group who contracted COVID, the ones who received the vaccine had more symptoms than those with the placebo while drug user group that did not contract COVID showed no difference between those who were vaccinated and not vaccinated.

## Numerical

```{r}
table1 <- xtabs(~infected + treatment,data=dataD)
rowPerc(table1)
```

```{r}
colPerc(table1)
```

The data from this table shows that from the druggies group that were infected with COVID 19, The unvaccinated 11.67% of people experienced symptoms while the vaccinated 88.33% of the druggies group experienced symptoms. It is also showing that within the druggies placebo group, 0.4% were infected with COVID symptoms , while 3.06% of the vaccinated group contracted COVID symptoms.


## Inferential Results

```{r}
chisq.test(table1)
```

```{r}
chisqtestGC(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Here this chart shows the exact same p-value as the previous chart. In similar fashion, we will reject the null hypothesis seeing as though the p-value is zero. We will therefore conclude that there is a significant association between treatment type and infection status. Based on the Fisher exact test, we got a p-value of 0.000000000232, which suggest the same as the Chi-squared test.

# Overall Results and Conclusions
In conclusion, the effectiveness of the COVID vaccine varies significantly across different population subgroups. The vaccine showed a significant reduction in the percentage of females contracting COVID-19 compared to those who received the placebo. Both the Chi-squared test and Fisher exact test confirmed this result (p < 0.001). Within the LGBTQ subgroup, individuals who received the vaccine showed a higher percentage of COVID-19 symptoms compared to those who received the placebo. Both statistical tests confirmed this unexpected result (p < 0.001). Similar to the LGBTQ subgroup, druggies who got the vaccine showed a higher likelihood of COVID-19 infection (p < 0.001).Therefore, based on the statistical tests conducted for each subgroup, we reject the null hypothesis that the COVID vaccine has no difference in effectiveness on people in different subgroups.
