---
title: "Covid91 vaccine study Final2023"
author: "Matt Diaz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: " April 3, 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_o$: There is no difference in the effectiveness of the covid 921 vaccine in females.

$h_a$: There is a difference in the effectiveness of the civid 921 vaccine in females.

## Methods

The variables in this study: Males-categorical, Females-categorical, LGBTQ-categorical, Druggies, categorical. We will be using the Cat~Cat helper file. 

## Graphical Descriptive Results

#Females

```{r}
dataF = filter(FinalData,sex=="F")
barchartGC(~infected + treatment,data=dataF)
```

```{r}
barchartGC(~infected + treatment, data=dataF, type= "percent")
```

Based on the barcharts presented, there is significant percentage difference in the vaccine and the placebo.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=dataF)
rowPerc(table1)
```

```{r}
colPerc(table1)
```

With the data presented, it seems as though the vaccine has a lower infection rate compared to the placebo. In addition, among those infected, the vaccine has a slightly better treatment outcome with there being a .67 difference between the placebo and vaccine. 


## Inferential Results

```{r}
chisq.test(table1)
```


```{r}
chisqtestGC(table1)
```

When looking at this data, these tests show us that the odds of one being infected with the Covid virus is approximately 1.49 times higher with those using the placebo than the vaccine. In addition, this test and chart hows us that the vaccine is more effective in reducing risk for contracting the Covid virus than the placebo.The p-va;ue presented is 3e-03, so we will reject the null hypothesis. 


# LGBTQ

## Introduction

$H_o$: There is no difference in effectiveness of the Covid vaccine between LGBTQ people.

$h_a$: There is a difference in effectiveness of the Covid vaccine between LGBTQ people.

## Methods

We will be using the Cat~Cat helper file. 

## Graphical Descriptive Results

```{r}
levels(FinalData$LGBTQ)
```

```{r}
dataG = filter(FinalData,LGBTQ=="gay")
barchartGC(~infected + treatment,data=dataG)
```

```{r}
barchartGC(~infected + treatment,data=dataG, type = "percent")
```

This chart shows that there is a significant difference between the placebo and the vaccine. Among those that were infected with covid, those who had the vaccine were infected more than those who had received the placebo. 

## Numerical Results 

```{r}
table1 <-xtabs(~infected + treatment, data=dataG)
rowPerc(table1)
```

```{r}
colPerc(table1)
```

The infection rate among those who had received the placebo is 15.38%. The infection rate among those who had received the vaccine is 0.66%. This data tell us that the vaccine is more effective in reducing the risk of Covid.


## Inferential Results

```{r}
chisq.test(table1)
```

```{r}
chisqtestGC(table1)
```

With the p-value being zero, we reject the null hypothesis and therefore conclude that there is a significant association between the treatment type and infection status. The treatment type, whether that be someone who was administered the vaccine or received the placebo, is related to getting infected with the virus. 


#Druggies

## Introduction

$H_o$: There is no difference in the effectiveness of COVID921 vaccine within druggies.

$h_a$:There is a difference in the effectiveness of COVID921 vaccine within druggies.

## Methods

We will be using the Cat~Cat helper file.

## Graphical Descriptive Results

```{r}
levels(FinalData$DrugUser)
```

```{r}
dataD = filter(FinalData,DrugUser=="yes")
barchartGC(~infected + treatment,data=dataD)
```

```{r}
barchartGC(~infected + treatment,data=dataD, type="percent")
```

This chart shows is that there is a significant difference between the drug use and between those who are "OK". 

## Numerical

```{r}
table1 <- xtabs(~infected + treatment,data=dataD)
rowPerc(table1)
```

```{r}
colPerc(table1)
```

Among all of the Covid cases that occurred among those who received the vaccine, there was a higher percentage of vaccine recipients who remained uninfected compared to the placebo group. 

## Inferential Results

```{r}
chisq.test(table1)
```

```{r}
chisqtestGC(table1)
```

Here this chart shows the exact same p-value as the previous chart. In similar fashion, we will reject the null hypothesis seeing as though the p-value is zero. We will therefore conclude that there is a significant association between treatment type and infection status. 

# Overall Results and Conclusions








